\textbf{Using the MSU Thesis Class}

### 2024-05-15 docs/msu-thesis.tex
Added some explicit Overleaf instructions

\subsection{Customizing the look of things}
I have created a basic looking style for Chapter headings based on the |thatcher| style in |memoir|.  If you want to change these things, you may, using the appropriate |\renewcommand| commands.  More information can be found in the |memoir| documentation. The Thesis Office seems to require dotted leaders for the List of Figures/Tables in the Table of Contents, (despite the guide saying they are optional) even though they get typeset as Chapters.  If you wish your Chapter headings to have no leaders, add the following command just after the |\mainmatter| command.  To put the dots back for the Bibliography, repeat the same command right after the |\backmatter| command, but replace |\cftnodots| with |\cftdotsep|.  The default template leaves dots everywhere, which should appease the thesis office. 
\begin{quote}
\begin{lstlisting}
\addtocontents{toc}{%
   \protect\renewcommand{\protect\cftchapterdotsep}
   	{\cftnodots}}
\end{lstlisting}
\end{quote}
Despite there being no explicit formatting requirements for Chapter headings (other than that they start 1 in from the top of the page), the thesis office has been known to dislike many things, so if you do decide to change things, be prepared for rejection or a frustrating fight.
\subsection{Things to watch out for\label{watchout}}
\begin{itemize}
\item If you are having problems with the Thesis office about non-conforming formatting, please check first whether you have a local copy of the class instead of the latest version distributed via CTAN. The log file will reflect the date of the current version number. Remember if you are using the most current and updated TeXLive or MikTeX you will automatically have the latest version already and you should not need to install a local version. Overleaf users may need a local copy, since Overleaf is usually a year behind the TeXLive distribution schedule.
\item If your title is long and you want to put a newline (|\\|) into it to improve the spacing, you must precede the newline with |\protect|.
\item Make sure you pay attention to any ``Overfull hbox'' warnings when you are producing your final copy. You should make sure you eliminate all of them or else the thesis office is likely to reject your thesis. One common way to eliminate such warnings is to place the command |\sloppy| at the beginning of the paragraph that triggers the warning. You should only use this technique at the very end when you are completely done with everything.  Alternatively, rewording the text may also fix things. 
\item Remember to use the |\tableofcontents*| command so that your table of contents is not listed in the table of contents.
\item Remember to use the |{appendix}| environment for single appendices,  and the  |{appendices}| environment for multiple appendices. Do not use the standard LaTeX  |\appendix| command.
\item Remember also that \emph{all} appendices whether global or per-chapter should be |\chapter| commands in your source document. 
\item Remember to add the |\makebibliographypage| command before any bibliography in your document.
\end{itemize}
\subsection{Master’s Theses}
Master’s theses are sometimes relatively short, and you may think that you can just have |\section| commands with no chapters. This will \emph{not} work. The class is designed to use chapters (and as far as I can tell, the Graduate School also expects chapters).

>>>>

\subsection{Customizing the look of things}
I have created a basic looking style for Chapter headings based on the |thatcher| style in |memoir|.  If you want to change these things, you may, using the appropriate |\renewcommand| commands.  More information can be found in the |memoir| documentation. The Thesis Office seems to require dotted leaders for the List of Figures/Tables in the Table of Contents, (despite the guide saying they are optional) even though they get typeset as Chapters.  If you wish your Chapter headings to have no leaders, add the following command just after the |\mainmatter| command.  To put the dots back for the Bibliography, repeat the same command right after the |\backmatter| command, but replace |\cftnodots| with |\cftdotsep|.  The default template leaves dots everywhere, which should appease the thesis office. 
\begin{quote}
\begin{lstlisting}
\addtocontents{toc}{%
   \protect\renewcommand{\protect\cftchapterdotsep}
   	{\cftnodots}}
\end{lstlisting}
\end{quote}
Despite there being no explicit formatting requirements for Chapter headings (other than that they start 1 in from the top of the page), the thesis office has been known to dislike many things, so if you do decide to change things, be prepared for rejection or a frustrating fight.
\subsection{Things to watch out for\label{watchout}}
\begin{itemize}
\item If you are having problems with the Thesis office about non-conforming formatting, please check first whether you have a local copy of the class instead of the latest version distributed via CTAN. The log file will reflect the date of the current version number. Remember if you are using the most current and updated TeXLive or MikTeX you will automatically have the latest version already and you should not need to install a local version. Overleaf users may need a local copy, since Overleaf is usually a year behind the TeXLive distribution schedule.
\item If your title is long and you want to put a newline (|\\|) into it to improve the spacing, you must precede the newline with |\protect|.
\item Make sure you pay attention to any ``Overfull hbox'' warnings when you are producing your final copy. You should make sure you eliminate all of them or else the thesis office is likely to reject your thesis. One common way to eliminate such warnings is to place the command |\sloppy| at the beginning of the paragraph that triggers the warning. You should only use this technique at the very end when you are completely done with everything.  Alternatively, rewording the text may also fix things. 
\item Remember to use the |\tableofcontents*| command so that your table of contents is not listed in the table of contents.
\item Remember to use the |{appendix}| environment for single appendices,  and the  |{appendices}| environment for multiple appendices. Do not use the standard LaTeX  |\appendix| command.
\item Remember also that \emph{all} appendices whether global or per-chapter should be |\chapter| commands in your source document. 
\item Remember to add the |\makebibliographypage| command before any bibliography in your document.
\end{itemize}
\subsection{Master’s Theses}
Master’s theses are sometimes relatively short, and you may think that you can just have |\section| commands with no chapters. This will \emph{not} work. The class is designed to use chapters (and as far as I can tell, the Graduate School also expects chapters).
\subsection{Overleaf users}
If you're using Overleaf, it's usually best to upload a copy of the class to your project. Please remember that the file \emph{must} be called |msu-thesis.cls|. Naming it something different will cause it not to be found, and an older version will be loaded. Additionally, since Overleaf preserves the TeX Live year that you first opened the project, it's a good idea to change to the most recent TeX Live year that the site offers. You can do this by clicking on the Menu icon in the  editor and then selecting the most recent TeX Live version.
